---
title: "lab 2 244"
author: "Maya Szafraniec"
date: "1/16/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(janitor)
library(naniar)
library(VIM)
library(skimr)
library(ggfortify)
library(here)
```

## 1. get some data

```{r}
ca_pb<-read_csv(here("data","ca_pollution_burden.csv"))
ca_dem<-read_csv(here("data","ca_census_demographics_2010.csv"))
```

## 2. some data cleaning and wrangling

```{r}
ca_pb_nopct<-ca_pb %>% 
  clean_names() %>% 
  select(-contains("perc")) %>% 
  select(-contains("pctl")) %>% 
  select(-latitude,-longitude)
```

```{r}
ca_dem_clean<-ca_dem %>% 
  clean_names()
```

## 3. Polution burden indicators only

```{r}
ca_pb_subset<-ca_pb_nopct %>% 
  select(ozone:solid_waste,asthma:housing_burden)
```

Now lets do pca
use the prcomp() function

```{r, eval=FALSE}

#this gives an error because there are missings
pb_pca<-prcomp(ca_pb_subset,scale=TRUE)
```

Aside:exploring missingness!
```{r}
summary(ca_pb_subset)

gg_miss_var(ca_pb_subset) #gives a plot for the number of missings you have in each column

matrixplot(ca_pb_subset,sortby = "poverty") # like a heat map of missing ness, darkness of gray scale indicates the value if it is missnig then their is a red spot (dark is higher number)

#sort by so poverty is from high to low and missings are on top


```

remove by compleat case deletion, any observation with any na values

```{r}
ca_pb_nona<-ca_pb_subset %>% 
  drop_na()

summary(ca_pb_nona)

skim(ca_pb_nona) #v nice summary of the data

```

now, no NA values, cool set to do PCA, otherwise consider imputaiton 

```{r}
my_ca_pca<-prcomp(ca_pb_nona,scale=TRUE)

my_ca_pca
summary(my_ca_pca)

#biplot(my_ca_pca) #its gross, so make it cleaner

#actually use autoplot

#colour means dont show the location of points

my_biplot<-autoplot(my_ca_pca,
                    colour=NA,
                    loadings.label=TRUE,
                    loadings.label.size=3,
                    loading.lable.colour="black",
                    loading.label.repel =TRUE)

my_biplot


```

## 4. PCA for polution burden + some demographics

join together the ca_dem_clean with ca_pb_nopct

```{r}
ca_df<-ca_dem_clean %>% 
  inner_join(ca_pb_nopct,by=c("census_tract_number"="census_tract"))

#now deop NAs by complete case deletion
ca_nona<-ca_df %>% 
  drop_na()
```

maek a subset of ca_nona that only inculdes the following variables
-white_percent
-elderly_65_percent
-pm2_5
-pesticides
-traffic
-asthma
-cardiovasular_disease
-poverty

```{r}
my_sub<-ca_nona %>% 
  select(white_percent,elderly_65_percent,pm2_5,pesticides,traffic,asthma,cardiovascular_disease,poverty)
```

My PCA

```{r}
my_dem_pca<-prcomp(my_sub,scale=TRUE)

my_dem_biplot<-autoplot(my_dem_pca,
                    colour=NA,
                    loadings.label=TRUE,
                    loadings.label.size=3,
                    loading.lable.colour="black",
                    loading.label.repel =TRUE)+
  scale_y_continuous(limits=c(-0.05,0.05))+
  theme_minimal()

my_dem_biplot

```

